---
title: "490503902_Assignment"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(ggplot2)
library(skimr)
library(visdat)
```
# Executive summary
* Item 1
What are we analysing?

* Item 2

# Data Cleaning key points

**Lost a mark if they didn’t remove nonsensical values or otherwise didn’t clean the data before testing or didn’t talk about how the data was cleaned.**


Load data into R
Clean Data (1. Use one code chunk to perform all cleaning at once. 2. Write a couple of paragraphs describing rationale are all that's required)

```{r, message=FALSE, tidy=TRUE}
# The CITATION
raw = read_csv("class_survey.csv")
# changes all spaces into underscores and lowercases the whole string for column names
clean_df = raw %>% janitor::clean_names()
# Shorten the names of specific columns
colnames(clean_df) = stringr::str_replace(string = colnames(clean_df),
                                   pattern = "what_is_your_",
                                   replacement = ""
                                   )
colnames(clean_df) = stringr::str_replace(string = colnames(clean_df),
                                   pattern = "on_average_how_many_hours_per_week_did_you_",
                                   replacement = ""
                                   )
colnames(clean_df)[2] = "covid_tests"
colnames(clean_df)[4] = "postcode"
colnames(clean_df)[5] = "dentist"
colnames(clean_df)[6] = "university_work"
colnames(clean_df)[7] = "social_media"
colnames(clean_df)[8] = "dog_or_cat"
colnames(clean_df)[9] = "live_with_parents"
colnames(clean_df)[10] = "exercising"
colnames(clean_df)[12] = "asthma"
colnames(clean_df)[13] = "paid_work"
colnames(clean_df)[14] = "fav_season"
colnames(clean_df)[16] = "height"
colnames(clean_df)[17] = "floss_frequency"
colnames(clean_df)[18] = "glasses"
colnames(clean_df)[20] = "steak_preference"
colnames(clean_df)[21] = "stress_level"
## Change postcode to character
clean_df = clean_df %>% mutate(
  postcode = as.character(postcode)
)
table(clean_df["postcode"])
skimr::skim(clean_df)
# Change timestamp to POSIXct for easier handling
clean_df$timestamp <-  dmy_hms(clean_df$timestamp)

clean_df %>% filter(is.na(clean_df$gender))

# head(clean_df[1])
# m <- sapply(clean_df[1],as.POSIXct)
# m
```
```{r}
clean_df[1][1,]

# table(clean_df$timestamp)
# clean_df$timestamp <- as.Date(clean_df$timestamp,origin = "1970-01-01")
class(clean_df$timestamp)
# a <- table(clean_df$timestamp)
# a$
# aggregate(clean_df[2], by=clean_df[1],sum)
# clean_df %>% ggplot(aes(x=)) + geom_histogram()
# clean_df$timestamp <- as.POSIXlt(clean_df$timestamp)

# class(clean_df$timestamp)
# output
# "2014-10-03 22:30:27 CEST"                                         # Print updated data
# "2020-01-01" "2015-10-01" "2033-05-01"
```

In

Only the following columns were cleaned.

3 paragraphs to explain what was cleaned.


Every graph or plot needs to referenced in text and explained.

Perform analysis.


## Is this a random sample of DATA2002 students?

To answer this question, we need to define what is a random sample. A random sample is a 

equal probability of being chosen. This is not the case as the survey is geared more towards studious students who regularly check Edstem posts. Further, since the 
 
 
Yes, this is a sample with only 174 people recorded in the survey but is it truly random? The way the test data collected was through voluntary choice and was advertised... This suggests it was not random for the

Also, to be a truly random sample of the DATA2002 population, the sample must be a true representation of all the types of people from the DATA2002 population.

We are not sure if the sample population is representative of the true population.


## What are the potential biases? Which variables are most likely to be subjected to this bias?


## Are there any questions that needed improvement to generate useful data?

Shoe size is impossible to clean! Why?


## Does the number of COVID tests follow a Poisson distribution?
**Do a chi-square hypothesis test**
To test this question, we will use a chi-square...

$H_0:$ The number of COVID tests does not follow a Poisson distribution
$H_1:$ The number of COVID tests does follow a Poisson distribution

**Assumptions**:

**Decision**:


## Perform two other hypothesis tests. Give some rationale for why you selected these hypothesis tests and interpret the results. Be sure to mention any limitations in the data that may impact your findings.

## Cleaning survey data

### Read in the data




