---
title: "lab1c"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quick quiz

Q1.1. B). We want to see if 2 categorical variables are independent.
Q1.2. C)
Q1.3. Yes, it does look like there's a relationship between age group and violence viewing.
Q1.4. Yes or no. Whatever way, you can argue.

## Group work
```{r}
library("gt")
library("")
survey = readr::read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vTf8eDSN_2QbMTyVvO5bdYKZSEEF2bufDdYnLnL-TsR8LM-6x-xu1cxmDMohlbLrkMJn9DE7EG7pg5P/pub?gid=1724783278&single=true&output=csv")
# str(survey)
colnames(survey) %>% tibble() %>% gt::gt()
```

## 3 Exercises
### 3.1. Personality type
$H_0:$ Probability distribution (proportions) of the business students = Probability distribution of social science students.
$H_1:$ Probability distribution (proportions) of the business students is not equal to probability distribution of social science students.

**Assumptions**: $e_{ij} \geq 5$ and independent observations.

**Decision**: As p-value > 0.05, we do not reject the null hypothesis.

```{r}
counts = c(41, 52, 46, 61, 58, 72, 75, 63, 80, 65)
c_mat = matrix(counts, nrow = 2, byrow = TRUE)
colnames(c_mat) = c("Open", "Conscientious", "Extrovert", "Agreeable", 
    "Neurotic")
rownames(c_mat) = c("Business", "Social Science")
```
To find chi square test (need to turn off Yates' correction, by using correct=FALSE),
```{r}
chisq.test(c_mat,correct=FALSE)
test = chisq.test(c_mat)
test$statistic
test$p.value
# use names to inspect a variable that contains a list.
names(test)
```
Now, to show it manually... No need to this manually for assignment.
We want to find the sum of population totals each one.
```{r}
# MARGIN = 1 means apply the sum FUNction 
# down rows. Alternative: rowSums(tab)
(yr = apply(c_mat, MARGIN = 1, FUN = sum))
```
```{r}
# MARGIN = 2 means apply the sum FUNction 
# across columns. Alternative: colSums(tab)
(yc = apply(c_mat, MARGIN = 2,FUN = sum))
```
```{r}
(yr.mat = matrix(yr, nrow = r, ncol = c, byrow = FALSE))
```

### 3.2. Shocking
```{r}
counts = c(12,5,4,9)
c_mat = matrix(counts,nrow=2,byrow = TRUE)

# i)
fisher.test(c_mat)
# ii & iii)
chisq.test(c_mat, correct = FALSE)
chisq.test(c_mat, correct = TRUE)
# iv)
chisq.test(c_mat, simulate.p.value = TRUE, B = 1000)

```

### 3.1. Personality type

